---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# forensicdatatoolkit

<!-- badges: start -->
<!-- badges: end -->

The goal of forensicdatatoolkit is to combine several tools created for investigating 
summary data and detect anomalies. This is not a tool for finding fraud. Fraud is a legal term.
These tools merely tell you if the underlying data is probable or not. It can never tell us
why the statistics are wrong. Someone could lie or someone could mistype a number. Only open data
can tell us which it is.

## Installation

You can install the released version of forensicdatatoolkit from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("forensicdatatoolkit")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(forensicdatatoolkit)
## basic example code
```


What is in the package (citations in separate file)

- [x] Granularity-Related Inconsistency of Means (GRIM)
- [] Granularity-Related Inconsistency of Means Mapped to Error Repeats (GRIMMER)
- [] Sample Parameter Reconstruction via Iterative Techniques (SPRITE)


We can use these tools to investigate summary statistics. 
For example if we collect statistics from a set of papers we can investigate them.

```{r}
example_data <- data.frame(
  study = c(1, 1, 1, 2, 2, 2),
  description = rep(c("food acceptance", "darkness","likeability"),2),
  n = c(20, 20, 19, 40, 41, 40),
  means = c(3.5, 5.6, 2, 2.5, 8, 3),
  sds = c(1.34, 1.34, 2.56, 0.45, 2, 1),
  scale_min = rep(c(1,1,1),2),
  scale_max = rep(c(5, 9, 7))
)
example_data
```
```{r base r style}
example_data$mean_possible <- grim(example_data$means, example_data$n)
example_data[,c("study","description","mean_possible")]
```


```{r tidyverse style}
library(dplyr)
example_data %>% 
  mutate(mean_possible =grim(mean = means, n=n)) %>% 
  select(study, description, mean_possible)
```

CORVIDS ?

